---
title: "Geospatial Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---
#Setup Workspace
In this code block, we define directory locations, download "required" libraries, and importing data.   
```{r setup, message=F}
#Clear working dir
remove(list=ls())

#Define directories of interest
data_dir    <- "//storage.research.sesync.org/njones-data/Research Projects/BiodiversiTREE/spatial_analysis/II_work/"
output_dir  <- "//storage.research.sesync.org/njones-data/Research Projects/BiodiversiTREE/spatial_analysis/III_output/"
scratch_dir <- "C:\\ScratchWorkspace\\"
wbt_dir     <- "C:/WBT/whitebox_tools"

#Download required libraries
library('tidyr')
library('dplyr')
library('lubridate')
library('dygraphs')
library('xts')
library('raster')
library('sf')
library('tmap')

#define spatial projection
p<-c("+proj=utm +zone=18 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")

#gather data
dem<-raster(paste0(data_dir, "dem_serc"))
  writeRaster(dem, paste0(scratch_dir, "dem.tif"), overwrite=T)
  dem<-raster(paste0(scratch_dir, "dem.tif"))
  dem<-projectRaster(dem, crs = CRS(p))
pnts<-st_read(paste0(data_dir, "outlet_points_utm.shp"))
  pnts<-st_transform(pnts, crs = p)
plots<-st_read(paste0(data_dir,"plots_2015.shp"))
  plots<-st_transform(plots, crs=p)
soils<-st_read(paste0(data_dir,"SSURGO/spatial/soilmu_a_aoi.shp"))
  soils_data<-read.csv(paste0(data_dir,"soils_data.csv"))
  soils_data<-soils_data %>% select(MUKEY=Mapunit_Key,
                                    map_unit=Mapunit_Symbol,
                                    hydro=Hydrologic_Group_Dominant_Conditions)
  soils$MUKEY<-as.numeric(paste(soils$MUKEY))
  soils<-left_join(soils, soils_data, by='MUKEY')  
  soils<-st_transform(soils, crs= p)
  soils<-st_crop(soils, y=st_bbox(plots))
  soils$MUKEY<-paste(soils$MUKEY)

#Save image
save.image(paste0(data_dir,'setup.RData'))  
```


#Watershed Delineation
In this code chunk, we use Whitebox GAT tools [WBT] to delineate the watersheds. To do this, we first need to download the WBT exacutable and place it in the C:\\ directory.  We also need to create a scratch workspace directory.  [Note, both of these locations were defined in the 'setup' chunk above.]  To use WBT, export the raster from the R environment into the workspace directory. Then, execute WBT commands using system calls. Once WBT tasks are complete, then import updated raster file back into the R enviornment. Easy peasy. 
```{r watershed_delineation, message=F}
#Write DEM to workspace
dem<-na.omit(dem)
dem<-dem+1000
writeRaster(dem, paste0(scratch_dir,"dem.tif"), overwrite=T)

#Write points to workspace
st_write(pnts, paste0(scratch_dir,"pnts.shp"), delete_layer = T)

#Smooth DEM with a Gaussian Filter
system(paste(paste(wbt_dir), 
             "-r=GaussianFilter", 
             paste0("--wd=",scratch_dir),
             "-i='dem.tif'", 
             "-o='dem_filter'",
             "--sigma=3"))

#Fill "single cell" depressions
system(paste(paste(wbt_dir),
              "-r=FillSingleCellPits",
              paste0("--wd=",scratch_dir),
              "--dem='dem_filter.tif'",
              "-o='dem_breach_minor.tif'"))

#Breach larger depressions
system(paste(paste(wbt_dir), 
             "-r=BreachDepressions", 
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_minor.tif'", 
             "-o='dem_breach_major.tif'"))

#Create Flow Accumulation Raster
system(paste(paste(wbt_dir), 
             "-r=D8FlowAccumulation", 
             "--out_type='cells'",
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'", 
             "-o='fac.tif'"))

#Create Stream Raster [fac>1000]
fac<-raster(paste0(scratch_dir,"fac.tif"))
fac[fac<1000]<-NA
fac<-fac*0+1
fac@crs<-CRS(p)
writeRaster(fac,paste0(scratch_dir,"flowgrid.tiff"), overwrite=T)
 
#Run flow direction 
system(paste(paste(wbt_dir),
             "-r=D8Pointer",
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'",
             "-o='fdr.tif'",
             "--out_type=sca"))

#Create pour pnt raster
system(paste(paste(wbt_dir),
             "-r=VectorPointsToRaster",
             paste0("--wd=",scratch_dir),
             "-i='pnts.shp'",
             "--field=FidActl",
             "-o=pp.tif",
             "--assign=min",
             "--nodata",
             "--base=dem.tif"))

#Jenson Snap Pour point
system(paste(paste(wbt_dir),
             "-r=JensonSnapPourPoints",
             paste0("--wd=",scratch_dir),
             "--pour_pts='pp.tif'",
             "--streams='flowgrid.tif'",
             "-o='pp_snap.tif",
             "--snap_dist=1000"
))

#Convert back to point file
snapgrid<-raster(paste0(scratch_dir,"pp_snap.tif"))
snappnts<-data.frame(rasterToPoints(snapgrid, fun=function(x){x>0}))
snappnts<-st_as_sf(snappnts,
            coords=c("x","y"),
            crs=paste(dem@crs))

#Export point back to scratch workspace
st_write(snappnts, paste0(scratch_dir,"snap.shp"), delete_layer = T)
    
#Watershed Tool
system(paste(paste(wbt_dir), 
             "-r=Watershed", 
             paste0("--wd=",scratch_dir),
             "--d8_pntr='fdr.tif'",
             "--pour_pts='snap.shp'", 
             paste0("-o='watershed.tif'")))

#Import watershed raster into R environment
ws_grid<-raster(paste0(scratch_dir,"watershed.tif"), overwrite=T)

#Define individual watersheds
ws110<-ws_grid
  ws110[ws110==2]<-NA
ws109<-ws_grid
  ws109[ws109==1]<-NA
  ws109[ws109==2]<-1
  
#convert to shapes
watersheds<-rasterToPolygons(ws_grid, dissolve = T)
  watersheds<-st_as_sf(watersheds, crs=p)
  
#Export shapes
writeRaster(ws_grid , paste0(data_dir, "watersheds.tif"), overwrite=T)
st_write(watersheds, paste0(output_dir, "watersheds.shp"), delete_layer = T)
  
#Save image
save.image(paste0(data_dir,'watershed_delineation.RData'))
```

#Plot Charactersitics
In this code chunk, the goal is to tabulate relevant topographic and soil propoerties for each plot

```{r twi_analysis, message=F}
#Soils Data~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
plots<-st_join(plots, soils, largest=T)

#Topographic Wetness index~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#Export dem to scratchworkspace
writeRaster(dem, paste0(scratch_dir, "dem.tif"), overwrite=T)

#Smooth DEM with a Gaussian Filter
system(paste(paste(wbt_dir), 
             "-r=GaussianFilter", 
             paste0("--wd=",scratch_dir),
             "-i='dem.tif'", 
             "-o='dem_filter'",
             "--sigma=3"))

#Fill "single cell" depressions
system(paste(paste(wbt_dir),
              "-r=FillSingleCellPits",
              paste0("--wd=",scratch_dir),
              "--dem='dem_filter.tif'",
              "-o='dem_breach_minor.tif'"))

#Breach larger depressions
system(paste(paste(wbt_dir), 
             "-r=BreachDepressions", 
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_minor.tif'", 
             "-o='dem_breach_major.tif'"))

#Create Flow Accumulation Raster
system(paste(paste(wbt_dir), 
             "-r=D8FlowAccumulation", 
             "--out_type='cells'",
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'", 
             "-o='fac.tif'"))

#Estimate slope
system(paste(paste(wbt_dir), 
             "-r=Slope",
             "-i=dem.tif",
             "-o=slope.tif",
             paste0("--wd=",scratch_dir)))

#Eatimate TWI
system(paste(paste(wbt_dir), 
             paste0("--wd=",scratch_dir),
             "-r=WetnessIndex",
             "--slope=slope.tif",
             "--sca=fac.tif",
             "-o=twi.tif"))

#Read into R Environmnet
e<-extent(plots)
twi<-raster(paste0(scratch_dir,"twi.tif"))
twi<-crop(twi, c(e[1],e[3],e[2],e[4]))

#Create function to estimate max TWI based on plot
fun<-function(n){
  #select plot
  site<-plots[n,]
  
  #Crop twi raster to area
  twi<-mask(twi, site)
  
  #Estimate mean twi
  mean<-cellStats(twi,mean)
  max<-cellStats(twi,max)
  
  #export
  c(site$plot, mean, max)
}

#Estimate twi and merge with plots
twi<-lapply(seq(1,dim(plots)[1]), fun)
twi<-data.frame(do.call(rbind, twi))
  colnames(twi)<-c("Plot_1", "twi_mean", "twi_max")
plots<-left_join(plots, twi, by='Plot_1')

#Export shapefile
st_write(plots, paste0(output_dir,"plots.shp"), delete_layer = T)

#save image
save.image(paste0(data_dir,'plot_characterization.RData'))

#Plot!
tmap_mode('plot')
tm1<-tm_shape(plots) +
  tm_polygons("twi_mean", palette = "Blues", style = 'quantile', breaks=8) +
  tm_layout(frame = F, legend.format = list(digits=1))
tm2<-tm_shape(plots) +
  tm_polygons("twi_max", palette = "Blues", style = 'quantile', breaks=8) +
  tm_layout(frame = F, legend.format = list(digits=1))
tm3<-tm_shape(plots) +
  tm_polygons("hydro", palette = "RdYlBu", style = 'quantile', breaks=2) +
  tm_layout(frame = F)
plots$map_unit<-paste(plots$map_unit)
tm4<-tm_shape(plots) +
  tm_polygons("map_unit", palette = "Set1", style = 'quantile', breaks=8) +
  tm_layout(frame = F)
tmap_arrange(tm1, tm2, tm3, tm4)

```






