---
title: "R Notebook"
output: html_notebook
---
#Setup Workspace
In this code block, we define directory locations, download "required" libraries, and importing data.   
```{r setup, message=F}
#Clear working dir
remove(list=ls())

#Define directories of interest
data_dir    <- "//storage.research.sesync.org/njones-data/Research Projects/BiodiversiTREE/spatial_analysis/II_work/"
scratch_dir <- "C:\\ScratchWorkspace\\"
wbt_dir     <- "C:/WBT/whitebox_tools"

#Download required libraries
library('tidyr')
library('dplyr')
library('lubridate')
library('dygraphs')
library('xts')
library('raster')
library('sf')

#define spatial projection
p<-c("+proj=utm +zone=18 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")

#gather data
dem<-raster(paste0(data_dir, "dem_serc"))
  writeRaster(dem, paste0(scratch_dir, "dem.tif"), overwrite=T)
  dem<-raster(paste0(scratch_dir, "dem.tif"))
  dem<-projectRaster(dem, crs = CRS(p))
pnts<-st_read(paste0(data_dir, "outlet_points_utm.shp"))
  pnts<-st_transform(pnts, crs = p)

#Save image
save.image(paste0(data_dir,'setup.RData'))  
```


#Watershed Delineation
In this code chunk, we use Whitebox GAT tools [WBT] to delineate the watershed. To do this, we fiirst need to download the WBT exacutable and place it in the C:\\ directory.  We also need to create a workspace directory.  [Note, both of these locations were defined in the 'setup' chunk above.]  To use WBT, export the raster from the R environment into the workspace directory. Then, execute WBT commands using system calls. Once WBT tasks are complete, then import updated raster file back into the R enviornment. Easy peasy. 
```{r}
#Write DEM to workspace
dem<-na.omit(dem)
dem<-dem+1000
writeRaster(dem, paste0(scratch_dir,"dem.tif"), overwrite=T)

#Write points to workspace
st_write(pnts, paste0(scratch_dir,"pnts.shp"), delete_layer = T)

#Smooth DEM with a Gaussian Filter
system(paste(paste(wbt_dir), 
             "-r=GaussianFilter", 
             paste0("--wd=",scratch_dir),
             "-i='dem.tif'", 
             "-o='dem_filter'",
             "--sigma=3"))

#Fill "single cell" depressions
system(paste(paste(wbt_dir),
              "-r=FillSingleCellPits",
              paste0("--wd=",scratch_dir),
              "--dem='dem_filter.tif'",
              "-o='dem_breach_minor.tif'"))

#Breach larger depressions
system(paste(paste(wbt_dir), 
             "-r=BreachDepressions", 
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_minor.tif'", 
             "-o='dem_breach_major.tif'"))

#Create Flow Accumulation Raster
system(paste(paste(wbt_dir), 
             "-r=D8FlowAccumulation", 
             "--out_type='cells'",
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'", 
             "-o='fac.tif'"))

#Create Stream Raster [fac>1000]
fac<-raster(paste0(scratch_dir,"fac.tif"))
fac[fac<1000]<-NA
fac<-fac*0+1
fac@crs<-CRS(p)
writeRaster(fac,paste0(scratch_dir,"flowgrid.tiff"), overwrite=T)
 
#Run flow direction 
system(paste(paste(wbt_dir),
             "-r=D8Pointer",
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'",
             "-o='fdr.tif'",
             "--out_type=sca"))

#Create pour pnt raster
system(paste(paste(wbt_dir),
             "-r=VectorPointsToRaster",
             paste0("--wd=",scratch_dir),
             "-i='pnts.shp'",
             "--field=FidActl",
             "-o=pp.tif",
             "--assign=min",
             "--nodata",
             "--base=dem.tif"))

#Jenson Snap Pour point
system(paste(paste(wbt_dir),
             "-r=JensonSnapPourPoints",
             paste0("--wd=",scratch_dir),
             "--pour_pts='pp.tif'",
             "--streams='flowgrid.tif'",
             "-o='pp_snap.tif",
             "--snap_dist=1000"
))

#Convert back to point file
snapgrid<-raster(paste0(scratch_dir,"pp_snap.tif"))
snappnts<-data.frame(rasterToPoints(snapgrid, fun=function(x){x>0}))
snappnts<-st_as_sf(snappnts,
            coords=c("x","y"),
            crs=paste(dem@crs))

#Export point back to scratch workspace
st_write(snappnts, paste0(scratch_dir,"snap.shp"), delete_layer = T)
    
#Watershed Tool
system(paste(paste(wbt_dir), 
             "-r=Watershed", 
             paste0("--wd=",scratch_dir),
             "--d8_pntr='fdr.tif'",
             "--pour_pts='snap.shp'", 
             paste0("-o='watershed.tif'")))

#Import watershed raster into R environment
ws_grid<-raster(paste0(scratch_dir,"watershed.tif"), overwrite=T)

#Define individual watersheds
ws110<-ws_grid
  ws110[ws110==2]<-NA
ws109<-ws_grid
  ws109[ws109==1]<-NA
  ws109[ws109==2]<-1
  
#convert to shapes
watersheds<-rasterToPolygons(ws_grid, dissolve = T)
  watersheds<-st_as_sf(watersheds, crs=p)
  
#Export shapes
writeRaster(ws_grid , paste0(data_dir, "watersheds.tif"), overwrite=T)
st_write(watersheds, paste0(data_dir, "watersheds.shp"), delete_layer = T)
  
#Save image
save.image(paste0(data_dir,'watershed_delineation.RData'))

```

