---
title: "Paired Wateshed Exploratory Analysis"
output:
  html_document:
    df_print: paged
    toc: yes
      toc_float: yes
---

The goal of this workbook is to explore impact of landuse change on downstream hydrology at the SERC biodiversiTREE site near Edgewater, MD. Here, we utilize longterm flow data from two "paired" watersheds (watersheds 109 and 110). Watershed 109 ecompasses most of the biodiversiTREE experiment, where landuse changed from agricultural [e.g., corn field] to forested in 2013, and watershed 110 is a forested watershed that serves as a control.  

#Setup Workspace
This code chunk organizes the workspace. 
```{r setup, message=F}
#Define directories of interest
data_dir<-"/nfs/njones-data/Research Projects/BiodiversiTREE/data/flow_data/"

#Download required libraries
library('tidyr')
library('dplyr')
library('lubridate')
library('dygraphs')
library('xts')

#Download data
ws109<-read.csv(paste0(data_dir,'dflo109_07_18.csv'))
ws110<-read.csv(paste0(data_dir,'dflo110_07_18.csv'))

#Convert date format to POSIX
ws109$DAY<-as.POSIXct(ws109$DAY, format="%d-%b-%y")
ws110$DAY<-as.POSIXct(ws110$DAY, format="%d-%b-%y")

#Convert to cfs [from ft3/d]
ws109 %<>% mutate(cfs= FLOW__CF/86400) %>% select(DAY,cfs)
ws110 %<>% mutate(cfs= FLOW__CF/86400) %>% select(DAY,cfs)

#Combine flow data
df<-left_join(ws109, ws110, by="DAY")
df<-na.omit(df)
colnames(df)<-c("day", "Q109", "Q110")
remove(list=ls()[ls()!='df'])
```

#Plots of Flow Data
Below are interactive plots of flow data from each watershed. Click and drag mouse to zoom to time period of interest, double click to zoom back out to entire extent, and use the sliding bar at the bottom to pan accross the time period. 

##Watershed 109
```{r plot1, echo=F}
df_xts<-xts(data.frame(day  = df$day, 
                       flow = df$Q109), 
            order.by=df$day)
dygraph(df_xts, main="Watershed 109", ylab="Flow [cfs]") %>%
    dyRangeSelector() %>%
    dyAxis("y", label = "Flow [cfs]", valueRange = c(0, 3)) %>%
    dyOptions(strokeWidth = 1.5) %>%
    dyOptions(labelsUTC = TRUE) %>%
    dyHighlight(highlightCircleSize = 5,
                highlightSeriesBackgroundAlpha = 0.2,
                hideOnMouseOut = FALSE)

```

##Watershed 110
```{r plot2, echo=F}
df_xts<-xts(data.frame(day  = df$day, 
                       flow = df$Q110), 
            order.by=df$day)
dygraph(df_xts, main="Watershed 110", ylab="Flow [cfs]") %>%
    dyRangeSelector() %>%
    dyAxis("y", label = "Flow [cfs]", valueRange = c(0, 3)) %>%
    dyOptions(strokeWidth = 1.5) %>%
    dyOptions(labelsUTC = TRUE) %>%
    dyHighlight(highlightCircleSize = 5,
                highlightSeriesBackgroundAlpha = 0.2,
                hideOnMouseOut = FALSE)

```

#Residual Analysis
##Annual Water Volume Export
As a first step, we examine water export [volume] from the two watersheds. 
```{r resid_daily, echo=F}
#Filter outliers
df<- df %>% filter(abs(df$Q109-df$Q110)<10)

#aggregate to annual scale
year<-df %>% 
  mutate(month=month(day), 
         year =year(day)) %>%
  group_by(year) %>%
  summarise(vol_109 = sum(Q109), 
            vol_110 = sum(Q110)) %>%
  mutate(diff=vol_109-vol_110) %>%
  arrange(year)

#Estimate residuals
model<-lm(vol_109~vol_110, data=year)
year$res<-model$residuals

#Plot Annual Flow [Total Volume]
plot(year$year, year$vol_109, pch=19, col="grey30")
plot(year$year, year$vol_110, pch=19, col="grey30")
points(year$year, year$vol_109)


plot(year$vol_109~year$vol_110)
points(year$vol_109[year$year>2012]~year$vol_110[year$year>2012], pch=19, col="red")
abline(model)
abline(a=1, b=1)
  
plot(year$year, year$res, type="n", col="grey30")
  abline(v=2013)
  lines(l_model, col="red")



#Monthly differences in flow
# month<-df %>% 
#   mutate(month=month(day), 
#          year =year(day)) %>%
#   group_by(month,year) %>%
#   summarise(vol_109 = sum(Q109), 
#             vol_110 = sum(Q110)) %>%
#   mutate(diff=vol_109-vol_110) %>%
#   filter(abs(diff)<100000) %>%
#   arrange(year, month)
# plot((month$year+month$month/12), month$diff, type="l", col="grey30")
#   abline(v=2013)
# 

```




























